---
title: "Analysis of Stack Overflow Q&A"
author: "Semih Barutcu"
date: "12/11/2020"
output: beamer_presentation
---

```{r warning=F, message=F}
library(pacman)

p_load(dplyr, tidytext, ggplot2, wordcloud, skimr, stringr, lubridate)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

- Exploring the relationship between questions and tags.

- Using Text Mining tools

- Predicting tags from the question bodies.

- Resource Book: Text Mining with R by Julia Silge and David Robinson. <https://www.tidytextmining.com/>

## About The Dataset

- A kaggle dataset: StackSample: 10% of Stack Overflow Q&A. <https://www.kaggle.com/stackoverflow/stacksample>.

- Dataset with the text of 10% of questions and answers from the Stack Overflow.

- Used Questions and Tags datasets.

## The Most Popular Tags

```{r, message=F}
tags %>%  
  group_by(tag) %>% 
  summarize(n = n()) %>% 
  top_n(15) %>% 
  ggplot(aes(tag, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

## The Most Popular Words

```{r, message=F}
q_small %>%  
  group_by(word) %>% 
  anti_join(stop_words, by = "word") %>% 
  summarize(n = n(), .groups = 'drop') %>%
  top_n(20) %>% 
  ggplot(aes(word, n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip()
```

## Popularity of Programming Languages by Year

```{r}
questions %>% select(Id, CreationDate) %>%
  left_join(tags, by = "Id") %>%
  filter(tag %in% c( "ruby", "python", "php", "javascript", "java", "c")) %>%
  ggplot(aes(year(CreationDate))) +
  geom_bar() +
  facet_wrap(~tag) +
  scale_x_continuous(breaks = 2007:2016, ) + 
  theme(axis.text.x = element_text(angle = 45))

```

## Average Scores of the Selected Languages

```{r message=F}
questions %>% select(Id, Score) %>%
  left_join(tags, by = "Id") %>%
  filter(tag %in% c( "ruby", "python", "php", "javascript", "java", "c")) %>%
  group_by(tag) %>%
  summarise(Avg = round(mean(Score), 2), Min = min(Score), Max = max(Score), Q_025 = quantile(Score, 0.025), Q_975 = quantile(Score, 0.975) ) %>%
  arrange(desc(Avg))
```


## Sentiment Analysis

- Used AFINN lexicon.

```{r}
q_small %>% 
  select(Id, word) %>%
  inner_join(get_sentiments("afinn"), by = "word") %>%
  group_by(Id) %>%
  summarise(sentiment = sum(value), .groups = 'drop') %>%
  head()
```



## Top 10 Positive and Negative Words

```{r}
q_small_afinn_count %>%
  group_by(sentiment) %>%
  top_n(10, n) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
  x = NULL) +
  coord_flip()
```

## Term Frequency

```{r message=F, warning=F}
q_words %>%
  left_join(tags, by = "Id") %>%
  filter(tag %in% c( "ruby", "python", "php", "javascript", "java", "c")) %>%
  select(-Id,-n,-total)
  group_by(tag) %>%
  bind_tf_idf(word, tag, n) %>% 
  arrange(desc(tf_idf)) 
```

## The Words with Highest tf_idf Scores of Selected Languages

```{r message=F, warning=F}
q_words %>%
  left_join(tags, by = "Id") %>%
  filter(tag %in% c( "ruby", "python", "php", "javascript", "java", "c")) %>%
  group_by(tag) %>% 
  bind_tf_idf(word, tag, n) %>%
  select(-Id) %>%
  arrange(desc(tf_idf)) %>%
  top_n(5) %>%
  ggplot(aes(word, tf_idf, fill = tag)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~tag, scales = "free_y")+
  coord_flip() + 
  theme(axis.text.x = element_text(angle = 30))
```

## Conclusion
























